\section{Broader Impacts}

\paragraph{Improving Software System Reliability:} A key element of
this proposal's approach is to focus on realistically deployable
techniques.  We initially aim
for strong integration with NASA's FPrime~\cite{fprime,fprimerepo}
open source
flight software architecture and platform; PI Groce is already in
discussion with engineers at NASA's Jet Propulsion Laboratory, and
engaged in producing tests for the FPrime autocoder using DeepState,
and has previously sent summer students to work with NASA's FPrime team.
This integration will allow our
methods to be applied to CubeSat missions, instruments, and other flight software
systems, leading to improved reliability for low-budget space-based
scientific efforts.  We expect, in the long run, that our approaches
will lead to more reliable and robust development in many embedded and
cyberphysical systems domains.  Moreover, advances in precondition
handling in fuzzing, extension of DeepState to use model checking
backends for full verification of some kinds of code, and empirical
evidence on the value of domain-specific mutation operators will apply
to software bug-finding efforts well beyond the domain of
framework-based embedded systems.

\paragraph{Education and Outreach:}
The proposed research yields several opportunities for enhancing CS
education, recruiting new CS majors, and retaining CS students.  PI Groce will work with the NAU Student ACM Chapter to
present a series of ``excursions in testing'' that introduce automated
testing to students, using DeepState to find bugs in real world code,
including code from media player libraries.  The
work of Guzdial \cite{Guzdial} has shown that media computation is an
effective way to both recruit and retain students in computer science. Groce also teaches
a class on automated testing of embedded systems that can present
FPrime and project results.  Groce's prior
engagement in teaching and mentoring high school students (including
as summer students who have published their work in major venues) and working with
local technically-oriented 4H chapters, including in rural communities
(such as Williams, AZ) surrounding Flagstaff will also allow
preparation and presentation of versions of these excursions tailored
to high school students' level of knowledge.