\section{Research Plan}

\subsection{Automated Test Generation}
\input{framactodeepstate}

\subsection{Automated Specification Checking}
\input{modifies}

\subsection{Framework Aware Mutation Analysis}
\input{mutation}

\subsection{Integrating Software Model Checking}

\paragraph{Bounded Model Checking:}
\input{cbmcplan}

\paragraph{Explicit-State Model Checking:}
\input{spinplan}

\subsection{Case Studies} % for Ecological Monitoring and Control}
\label{sec:case-study}

The above approaches will be evaluated and designed in the context of
case studies based on actual FPrime models.  Initially, these will be
open source code, likely beginning with the FPrime system reference
from the FPrime community, and moving to actual flight projects as
available.  After initial work, we aim to collaborate with FPrime
users and team members at NASA/JPL to apply our prototypes to
in-progress missions in a non-disruptive fashion.  Finally, we will,
in the final months of the project, investigate developing a harness
transform mapping for a widely used framework with different
semantics and usage context, e.g. most likely AUTOSAR~\cite{AUTOSAR},
which is widely used in the automotive community, another important
domain for embedded systems with stringent correctness and reliability
requirements.

We will evaluate improvements to fuzzing (and other fault detection,
e.g., non-exhaustive model checking) approaches using both
traditional coverage and actual fault-based evaluations, and using
mutation testing.  There will be considerable synergy between
automation and extension of mutation for framework-based code and
evaluation of testing improvements for said code.

This project will support one graduate student, who will be involved
in all phases of the project, initially focusing on understanding
FPrime and DeepState, then working with PI Groce on the fundamental
harness generation work as a foundation for other aspects of the
project; we expect that initial work on the LLVM modifies checker and
framework-aligned mutation will be underway by the second year, with
model checker integration starting in the third year.  For this
project, we have budgeted summer pay for the student, since the amount
of software development the student will be responsible for is
unusually high.  If we are able to place the student in a summer JPL
position, this will be re-allocated to, e.g., hire undergraduates to work on implementation tasks,
recruited from PI Groce's upper-level classes, or for travel to work
with JPL.  